---
title: "ABCD.chronotype.descriptives"
output:
  html_document:
    df_print: paged
---

## Install packages #
```{r message=FALSE, warning=FALSE}
rm(list = ls())
options(stringsAsFactors = FALSE)

library(tidyverse)
library(naniar)
library(gtsummary)
library(ggpubr)
```

## Load pheontoye data #
```{r}
load("ABCD.Release.3.0.Merged.Data.Rda")
names(abcdData)[names(abcdData) == "subjectkey"] <- "src_subject_id"
```

## Load PRS #
```{r}
setwd("../data")
prs_chronotype <- read.delim("ABCD.PRS.Chronotype.Jones.txt",  header = TRUE, na.strings = c("", "NA"))
prs_chronotype <- dplyr::select(prs_chronotype, IID, SCORE)
colnames(prs_chronotype) <- c("src_subject_id", "PRS.CHRONOTYPE.JONES.NoMHC")
prs_chronotype$src_subject_id <- substr(prs_chronotype$src_subject_id, 11, 26)
```

## Merge phenotype and PRS data #
```{r}
abcdData <- merge(abcdData, prs_chronotype, by = c("src_subject_id"), all.x = TRUE)
```

## Filter based on eventname #
```{r}
abcdData <- filter(abcdData, eventname == "2_year_follow_up_y_arm_1")
```

## Keep participants with chronotype data #
```{r}
completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

abcdData_chronotype <- completeFun(abcdData, "mctq_msfsc_calc") 
abcdData_chronotype <- as.data.frame(abcdData_chronotype) 
```

## Keep participants with complete data #
```{r}
abcdData_chronotype <- abcdData_chronotype %>%
  replace_with_na(replace = list(sex = c(777,999)))

abcdData_chronotype <- abcdData_chronotype %>% filter(!is.na(PRS.CHRONOTYPE.JONES.NoMHC)) 
abcdData_chronotype <- abcdData_chronotype %>% filter(!is.na(mctq_msfsc_calc)) 
abcdData_chronotype <- abcdData_chronotype[abcdData_chronotype$ses_AvgIncomeParentEdu<=100,] 
abcdData_chronotype <- abcdData_chronotype %>% filter(!is.na(interview_age)) 
abcdData_chronotype <- abcdData_chronotype %>% filter(!is.na(ses_AvgIncomeParentEdu)) 
abcdData_chronotype <- abcdData_chronotype %>% filter(!is.na(sex)) 
abcdData_chronotype <- abcdData_chronotype %>% filter(!is.na(site_id_l)) 
```

## Rename levels of pubertal status #
```{r}
# Combine male/female
abcdData_chronotype <- abcdData_chronotype %>% mutate(pds_p_ss_all_category_2 = coalesce(pds_p_ss_male_category_2, pds_p_ss_female_category_2)) 
abcdData_chronotype$pds_p_ss_all_category_2 <- as.factor(abcdData_chronotype$pds_p_ss_all_category_2)
abcdData_chronotype <- abcdData_chronotype %>% filter(!is.na(pds_p_ss_all_category_2))
abcdData_chronotype %>% group_by(pds_p_ss_all_category_2) %>% tally()

## Define variable
abcdData_chronotype$pds_p_ss_all_category_2_num <- as.numeric(abcdData_chronotype$pds_p_ss_all_category_2)

## Rename
levels(abcdData_chronotype$pds_p_ss_all_category_2)
levels(abcdData_chronotype$pds_p_ss_all_category_2)[levels(abcdData_chronotype$pds_p_ss_all_category_2)=="1"] <- "Pre Puberty"
levels(abcdData_chronotype$pds_p_ss_all_category_2)[levels(abcdData_chronotype$pds_p_ss_all_category_2)=="2"] <- "Early Puberty"
levels(abcdData_chronotype$pds_p_ss_all_category_2)[levels(abcdData_chronotype$pds_p_ss_all_category_2)=="3"] <- "Mid Puberty"
levels(abcdData_chronotype$pds_p_ss_all_category_2)[levels(abcdData_chronotype$pds_p_ss_all_category_2)=="4"] <- "Late Puberty"
levels(abcdData_chronotype$pds_p_ss_all_category_2)[levels(abcdData_chronotype$pds_p_ss_all_category_2)=="5"] <- "Post Puberty"
levels(abcdData_chronotype$pds_p_ss_all_category_2)
```

## Fit 24 hr clock #
```{r}
my.plot <- hist.plot <- abcdData_chronotype %>%
    ggplot(aes(x=mctq_msfsc_calc, fill=mctq_msfsc_calc)) +
    geom_histogram(color="black", alpha=0.6) +
    labs(x='mctq_msfsc_calc') + 
    ggtitle('Chronotype (MFSfc)')
print(my.plot)

abcdData_chronotype$mctq_msfsc_calc.24hr <- ifelse(abcdData_chronotype$mctq_msfsc_calc >= 24, as.numeric(abcdData_chronotype$mctq_msfsc_calc)-24, abcdData_chronotype$mctq_msfsc_calc)

mean(abcdData_chronotype$mctq_msfsc_calc.24hr)
range(abcdData_chronotype$mctq_msfsc_calc.24hr)
hist(abcdData_chronotype$mctq_msfsc_calc.24hr, breaks = 50)

my.plot <- hist.plot <- abcdData_chronotype %>%
    ggplot(aes(x=mctq_msfsc_calc.24hr, fill=mctq_msfsc_calc.24hr)) +
    geom_histogram(color="black", alpha=0.6) +
    labs(x='mctq_msfsc_calc.24hr') + 
    ggtitle('Chronotype (MFSfc)')
print(my.plot)
```

### Limit to MSFc less than or equal to 12 #
```{r}
abcdData_chronotype <- abcdData_chronotype[abcdData_chronotype$mctq_msfsc_calc.24hr<=12,] 

my.plot <- hist.plot <- abcdData_chronotype %>%
    ggplot(aes(x=mctq_msfsc_calc.24hr, fill=mctq_msfsc_calc.24hr)) +
    geom_histogram(color="black", alpha=0.6) +
    labs(x='Chronotype (MSFc)', y = 'Count') 
print(my.plot)
```

## Rename levels of race/ethnicity #
```{r}
abcdData_chronotype$race_ethnicity <- as.factor(abcdData_chronotype$race_ethnicity)
levels(abcdData_chronotype$race_ethnicity)
levels(abcdData_chronotype$race_ethnicity)[levels(abcdData_chronotype$race_ethnicity)=="1"] <- "White"
levels(abcdData_chronotype$race_ethnicity)[levels(abcdData_chronotype$race_ethnicity)=="2"] <- "Black"
levels(abcdData_chronotype$race_ethnicity)[levels(abcdData_chronotype$race_ethnicity)=="3"] <- "Hispanic"
levels(abcdData_chronotype$race_ethnicity)[levels(abcdData_chronotype$race_ethnicity)=="4"] <- "Asian"
levels(abcdData_chronotype$race_ethnicity)[levels(abcdData_chronotype$race_ethnicity)=="5"] <- "Other"
levels(abcdData_chronotype$race_ethnicity)
```

## Plot pubertal status by sex #
```{r}
abcdData_chronotype$sex = as.factor(abcdData_chronotype$sex)
levels(abcdData_chronotype$sex)[levels(abcdData_chronotype$sex)=="M"] <- "Male"
levels(abcdData_chronotype$sex)[levels(abcdData_chronotype$sex)=="F"] <- "Female"

ggplot(abcdData_chronotype, aes(x=pds_p_ss_all_category_2, fill=sex, color=sex)) +
  geom_histogram(position="dodge", stat="count", alpha=0.6)
```


## Group by sex #
```{r}
male <- filter(abcdData_chronotype, sex == "Male") 
female <- filter(abcdData_chronotype, sex == "Female") 
```

## Table of descriptives #
```{r}
abcdData_chronotype$ancestry <- "European"

abcdData_chronotype %>%
  dplyr::select(sex, interview_age, ancestry, race_ethnicity, ses_AvgIncomeParentEdu, pds_p_ss_all_category_2, mctq_msfsc_calc.24hr, PRS.CHRONOTYPE.JONES.NoMHC) %>%
  rename("Age (Months)" = interview_age, "Socioeconomic Status (SES)" = ses_AvgIncomeParentEdu, "Pubertal Stage" = pds_p_ss_all_category_2, "MSFc" = mctq_msfsc_calc.24hr,  "Chronotype PGS" = PRS.CHRONOTYPE.JONES.NoMHC) %>%
  tbl_summary(digits = list(everything() ~ c(2)), by= sex, missing = "no") %>%
                add_p(pvalue_fun = function(x) style_pvalue(x, digits = 2)) %>% add_overall() %>% bold_labels() #%>%
  as_gt() %>% 
  gt::gtsave( 
    filename = "Table1.png"
  )
```

## MSFc range #
```{r}
range(abcdData_chronotype$mctq_msfsc_calc.24hr)
range(male$mctq_msfsc_calc.24hr)
range(female$mctq_msfsc_calc.24hr)
```

## Plot by age #
```{r}
my.plot <- ggplot(abcdData_chronotype, aes(y=mctq_msfsc_calc.24hr, x=interview_age, color=sex)) +
  geom_point(alpha=.3) + 
  geom_smooth() + stat_cor(method = "pearson", label.x = 155, size=3) +
  labs(y='Chronotype (MSFc)', x='Age (months)') +
  ggtitle('Chronotype vs Age by Sex')
print(my.plot) + theme_classic()
```

## MSFc residuals controlling for age #
```{r}
abcdData_chronotype_MSFCresid <- abcdData_chronotype
my.model <- lm (mctq_msfsc_calc.24hr ~ interview_age, data = abcdData_chronotype_MSFCresid)
summary(my.model)
abcdData_chronotype_MSFCresid$mctq_msfsc_calc.24hr_residualizedage <- my.model$resid
hist(abcdData_chronotype_MSFCresid$mctq_msfsc_calc.24hr_residualizedage)
```

# Group by pubertal stage #
```{r}
pre.puberty <- filter(abcdData_chronotype_MSFCresid, pds_p_ss_all_category_2 == "Pre Puberty")
early.puberty <- filter(abcdData_chronotype_MSFCresid, pds_p_ss_all_category_2 == "Early Puberty")
pre.early.puberty <- filter(abcdData_chronotype_MSFCresid, pds_p_ss_all_category_2 == "Pre Puberty" | pds_p_ss_all_category_2 == "Early Puberty" )
mid.puberty <- filter(abcdData_chronotype_MSFCresid, pds_p_ss_all_category_2 == "Mid Puberty")
late.puberty <- filter(abcdData_chronotype_MSFCresid, pds_p_ss_all_category_2 == "Late Puberty")
post.puberty <- filter(abcdData_chronotype_MSFCresid, pds_p_ss_all_category_2 == "Post Puberty")
late.post.puberty <- filter(abcdData_chronotype_MSFCresid, pds_p_ss_all_category_2 == "Late Puberty" | pds_p_ss_all_category_2 == "Post Puberty" )
```

## Compare MSFc between pubertal stages #
```{r}
wilcox.test(pre.puberty$mctq_msfsc_calc.24hr_residualizedage, late.post.puberty$mctq_msfsc_calc.24hr_residualizedage, alternative = "two.sided")
```

## Save data #
```{r}
save(abcdData_chronotype_MSFCresid,file="ABCD.Release.3.0.Merged.Data.Chronotype.Rda")
```
