---
title: "ABCD.chronotype.regressions"
output:
  html_document:
    df_print: paged
---

## Install packages #
```{r message=FALSE, warning=FALSE}
rm(list = ls())
options(stringsAsFactors = FALSE)

library(lme4)
library(lmerTest)
library(MuMIn)
library(sjPlot)
```

## Load pheontoye data #
```{r}
load("ABCD.Release.3.0.Merged.Data.Chronotype.Rda")
abcdData_chronotype <- abcdData_chronotype_MSFCresid
rm(abcdData_chronotype_MSFCresid)
```

## Group by sex and pubertal status #
```{r}
male <- filter(abcdData_chronotype, sex == "Male")
male.pre.puberty <- filter(male, pds_p_ss_all_category_2 == "Pre Puberty")
male.early.puberty <- filter(male, pds_p_ss_all_category_2 == "Early Puberty")
male.pre.early.puberty <- filter(male, pds_p_ss_all_category_2 == "Pre Puberty" | pds_p_ss_all_category_2 == "Early Puberty" )
male.mid.puberty <- filter(male, pds_p_ss_all_category_2 == "Mid Puberty")
male.late.puberty <- filter(male, pds_p_ss_all_category_2 == "Late Puberty")
male.post.puberty <- filter(male, pds_p_ss_all_category_2 == "Post Puberty")
male.late.post.puberty <- filter(male, pds_p_ss_all_category_2 == "Late Puberty" | pds_p_ss_all_category_2 == "Post Puberty" )

female <- filter(abcdData_chronotype, sex == "Female")
female.pre.puberty <- filter(female, pds_p_ss_all_category_2 == "Pre Puberty")
female.early.puberty <- filter(female, pds_p_ss_all_category_2 == "Early Puberty")
female.pre.early.puberty <- filter(female, pds_p_ss_all_category_2 == "Pre Puberty" | pds_p_ss_all_category_2 == "Early Puberty" )
female.mid.puberty <- filter(female, pds_p_ss_all_category_2 == "Mid Puberty")
female.late.puberty <- filter(female, pds_p_ss_all_category_2 == "Late Puberty")
female.post.puberty <- filter(female, pds_p_ss_all_category_2 == "Post Puberty")
female.late.post.puberty <- filter(female, pds_p_ss_all_category_2 == "Late Puberty" | pds_p_ss_all_category_2 == "Post Puberty" )

#pre.puberty <- filter(abcdData_chronotype, pds_p_ss_all_category_2 == "Pre Puberty")
#early.puberty <- filter(abcdData_chronotype, pds_p_ss_all_category_2 == "Early Puberty")
#pre.early.puberty <- filter(abcdData_chronotype, pds_p_ss_all_category_2 == "Pre Puberty" | pds_p_ss_all_category_2 == "Early Puberty" )
#mid.puberty <- filter(abcdData_chronotype, pds_p_ss_all_category_2 == "Mid Puberty")
#late.puberty <- filter(abcdData_chronotype, pds_p_ss_all_category_2 == "Late Puberty")
#post.puberty <- filter(abcdData_chronotype, pds_p_ss_all_category_2 == "Post Puberty")
#late.post.puberty <- filter(abcdData_chronotype, pds_p_ss_all_category_2 == "Late Puberty" | pds_p_ss_all_category_2 == "Post Puberty" )
```

## Regressions - all subjects #
```{r}
# Continuous chronotype PRS
my.model0 <- lmer(mctq_msfsc_calc.24hr ~ 
                   interview_age + ses_AvgIncomeParentEdu + pds_p_ss_all_category_2 + sex +
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
               data = abcdData_chronotype, na.action = na.omit, REML = FALSE)
print(summary(my.model0))
r.squaredGLMM(my.model0)

my.model1 <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                   interview_age + ses_AvgIncomeParentEdu + pds_p_ss_all_category_2 + sex +
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
               data = abcdData_chronotype, na.action = na.omit, REML = FALSE)
print(summary(my.model1))
r.squaredGLMM(my.model1)
anova(my.model0, my.model1)
```

## Regressions - by sex #
```{r}
# Male
my.model0 <- lmer(mctq_msfsc_calc.24hr ~ 
                 interview_age + ses_AvgIncomeParentEdu + pds_p_ss_all_category_2 +
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
                    data = male, na.action = na.omit, REML = FALSE)
print(summary(my.model0))
r.squaredGLMM(my.model0)

my.model1 <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                 interview_age + ses_AvgIncomeParentEdu + pds_p_ss_all_category_2 +
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
                    data = male, na.action = na.omit, REML = FALSE)
print(summary(my.model1))
r.squaredGLMM(my.model1) 
anova(my.model0, my.model1)

# Female
my.model0 <- lmer(mctq_msfsc_calc.24hr ~ 
                 interview_age + ses_AvgIncomeParentEdu + pds_p_ss_all_category_2 +
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l),  
                    data = female, na.action = na.omit, REML = FALSE)
print(summary(my.model0))
r.squaredGLMM(my.model0) 

my.model1 <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                 interview_age + ses_AvgIncomeParentEdu + pds_p_ss_all_category_2 +
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l),  
                    data = female, na.action = na.omit, REML = FALSE)
print(summary(my.model1))
r.squaredGLMM(my.model1) 
anova(my.model0, my.model1) 
```

## Interaction - sex #
```{r}
my.model <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE)*sex + 
                 interview_age + ses_AvgIncomeParentEdu + as.numeric(pds_p_ss_all_category_2) +
                 pc1 + pc2 + pc3 + pc4 +  pc5 + (1|site_id_l),
                    data = abcdData_chronotype, na.action = na.omit, REML = FALSE)
print(summary(my.model))
```

## Interaction - pubertal stage #
```{r}
my.model <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE)*as.numeric(pds_p_ss_all_category_2) + 
                 sex +
                 interview_age + ses_AvgIncomeParentEdu + 
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l),
                    data = abcdData_chronotype, na.action = na.omit, REML = FALSE)
print(summary(my.model))
```

## Interaction - sex + pubertal stage #
```{r}
my.model0 <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE)*sex + 
                as.numeric(pds_p_ss_all_category_2) +
                 interview_age + ses_AvgIncomeParentEdu +
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l),
                    data = abcdData_chronotype, na.action = na.omit, REML = FALSE)
print(summary(my.model0)) 

plot0 <- plot_model(my.model0, type = "pred", terms = c("PRS.CHRONOTYPE.JONES.NoMHC", "pds_p_ss_all_category_2", "sex"))
plot0$facet$params$nrow=2
plot0

my.model1 <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE)*sex + 
                scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE)*as.numeric(pds_p_ss_all_category_2) + 
                 interview_age + ses_AvgIncomeParentEdu +
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l),
                data = abcdData_chronotype, na.action = na.omit, REML = FALSE)
print(summary(my.model1)) 

anova(my.model0, my.model1) 

plot1 <- plot_model(my.model1, type = "pred", terms = c("PRS.CHRONOTYPE.JONES.NoMHC", "pds_p_ss_all_category_2", "sex"))
plot1$facet$params$nrow=2
plot1
```

## Regression by pubertal stage - male #
```{r}
# Pre puberty
my.model <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                 interview_age + ses_AvgIncomeParentEdu + 
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
                    data = male.pre.puberty, na.action = na.omit, REML = FALSE)
print(summary(my.model))    

# Early puberty
my.model <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                 interview_age + ses_AvgIncomeParentEdu + 
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
                    data = male.early.puberty, na.action = na.omit, REML = FALSE)
print(summary(my.model)) 

# Mid puberty
my.model <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                 interview_age + ses_AvgIncomeParentEdu + 
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
                    data = male.mid.puberty, na.action = na.omit, REML = FALSE)
print(summary(my.model)) 

# Late puberty
my.model <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                 interview_age + ses_AvgIncomeParentEdu + 
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
                    data = male.late.puberty, na.action = na.omit, REML = FALSE)
print(summary(my.model)) 
```

## Regression by pubertal stage - female #
```{r}
# Pre puberty
my.model <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                 interview_age + ses_AvgIncomeParentEdu + 
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
                    data = female.pre.puberty, na.action = na.omit, REML = FALSE)
print(summary(my.model)) 

# Early puberty
my.model <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                 interview_age + ses_AvgIncomeParentEdu + 
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
                    data = female.early.puberty, na.action = na.omit, REML = FALSE)
print(summary(my.model)) 

# Mid puberty
my.model <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                 interview_age + ses_AvgIncomeParentEdu + 
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
                    data = female.mid.puberty, na.action = na.omit, REML = FALSE)
print(summary(my.model)) 

# Late/post puberty
my.model <- lmer(mctq_msfsc_calc.24hr ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                 interview_age + ses_AvgIncomeParentEdu + 
                 pc1 + pc2 + pc3 + pc4 + pc5 + (1|site_id_l), 
                    data = female.late.post.puberty, na.action = na.omit, REML = FALSE)
print(summary(my.model)) 
```

## Prep data #
```{r}
abcdData_chronotype$bis_y_ss_bas_rr_plus1 <- (abcdData_chronotype$bis_y_ss_bas_rr +1)
which(colnames(abcdData_chronotype) == "bis_y_ss_bas_rr_plus1") #2537 #######

abcdData_chronotype$bis_y_ss_bas_drive_plus1 <- (abcdData_chronotype$bis_y_ss_bas_drive +1)
which(colnames(abcdData_chronotype) == "bis_y_ss_bas_drive_plus1") #2538 ########

abcdData_chronotype$pgbi_p_ss_score_plus1 <- (abcdData_chronotype$pgbi_p_ss_score +1)
which(colnames(abcdData_chronotype) == "pgbi_p_ss_score_plus1") #2539 ########

abcdData_chronotype$pps_y_ss_number_plus1 <- (abcdData_chronotype$pps_y_ss_number +1)
which(colnames(abcdData_chronotype) == "pps_y_ss_number_plus1") #2540 ########

abcdData_chronotype$pps_y_ss_severity_score_plus1 <- (abcdData_chronotype$pps_y_ss_severity_score +1)
which(colnames(abcdData_chronotype) == "pps_y_ss_severity_score_plus1") #2541 ##########
```

## MSFc PheWAS - baseline - all #
```{r}
datalist <- list(abcdData_chronotype)
names(datalist) <- c("abcdData_chronotype")

# Linear model
for (j in 1:length(datalist)) {
df.linear <- data.frame() 

  for (i in c(345, 346, 349:351, 353, 72, 336, 342)) {
    this_var = datalist[[j]][, i]
    my.model <- lm(as.matrix(this_var) ~ scale(mctq_msfsc_calc.24hr, center = TRUE, scale = TRUE) + 
                      interview_age + ses_AvgIncomeParentEdu + sex + pds_p_ss_all_category_2 +
                     pc1 + pc2 + pc3 + pc4 +  pc5 + pc6 + pc7 + pc8 + pc9 + pc10 +
                     site_id_l, 
                    data = datalist[[j]], na.action = na.omit)
df.linear <- rbind(df.linear, data.frame(data = names(datalist)[j],
                               ID = colnames(datalist[[j]])[i], 
                               beta_PRS = summary(my.model)$coefficients[2, 1],
                               stderr_PRS = summary(my.model)$coefficients[2, 2],
                               P_PRS = summary(my.model)$coefficients[2, 4]))
  }

# Gamma log-link
df.gamma <- data.frame() 
  for (i in c(10:16, 62:71, 73:81, 314:317, 319, 334:335, 337:338, 2538:2542)) {
    this_var = datalist[[j]][, i]
    my.model <- glm(as.matrix(this_var) ~ scale(mctq_msfsc_calc.24hr, center = TRUE, scale = TRUE) + 
                      interview_age + ses_AvgIncomeParentEdu + sex + pds_p_ss_all_category_2 +
                      pc1 + pc2 + pc3 + pc4 +  pc5 + pc6 + pc7 + pc8 + pc9 + pc10 + 
                      site_id_l, 
                    data = datalist[[j]], na.action = na.omit, family = Gamma(link = "log"))
df.gamma <- rbind(df.gamma, data.frame(data = names(datalist)[j],
                               ID = colnames(datalist[[j]])[i], 
                               beta_PRS = summary(my.model)$coefficients[2, 1],
                               stderr_PRS = summary(my.model)$coefficients[2, 2],
                               P_PRS = summary(my.model)$coefficients[2, 4]))
  }
  
df.all <- rbind(df.linear, df.gamma)
df.all$FDR_PRS <- p.adjust(df.by.ethnicity$P_PRS,'fdr')
print(df.all)

df.all[1, 2] = "NIH Toolbox Picture Vocabulary"
df.all[2, 2] = "NIH Toolbox Flanker"
df.all[3, 2] = "NIH Toolbox Pattern Comparison"
df.all[4, 2] = "NIH Toolbox Picture Memory"
df.all[5, 2] = "NIH Toolbox Reading"
df.all[6, 2] = "NIH Toolbox Crystalized"
df.all[7, 2] = "CBCL Total Problems"
df.all[8, 2] = "UPPS Sensation Seeking"
df.all[9, 2] = "BIS/BAS Fun Seeking"
df.all[10, 2] = "SDS Maintaining Sleep"
df.all[11, 2] = "SDS Sleep Breathing"
df.all[12, 2] = "SDS Disorders of Arousal"
df.all[13, 2] = "SDS Sleep-Wake Transition Disorders"
df.all[14, 2] = "SDS Excessive Somnolemce"
df.all[15, 2] = "SDS Sleep Hyperhydrosis"
df.all[16, 2] = "SDS Total Sleep Problems"
df.all[17, 2] = "CBCL Anx/Depress"
df.all[18, 2] = "CBCL Withdrawn/Depress"
df.all[19, 2] = "CBCL Somatic Complaints"
df.all[20, 2] = "CBCL Social Problems"
df.all[21, 2] = "CBCL Thought Problems"
df.all[22, 2] = "CBCL Attention Problems"
df.all[23, 2] = "CBCL Rule Breaking"
df.all[24, 2] = "CBCL Aggressive Behavior"
df.all[25, 2] = "CBCL Internalizing"
df.all[26, 2] = "CBCL Externalizing"
df.all[27, 2] = "CBCL Depression"
df.all[28, 2] = "CBCL Anxiety Disorders"
df.all[29, 2] = "CBCL Somatic Problems"
df.all[30, 2] = "CBCL ADHD"
df.all[31, 2] = "CBCL Oppositional Defiant"
df.all[32, 2] = "CBCL Conduct Problems"
df.all[33, 2] = "CBCL Sluggish Cog Tempp"
df.all[34, 2] = "CBCL OCD"
df.all[35, 2] = "CBCL Stress"
df.all[36, 2] = "BRIEF Attention"
df.all[37, 2] = "BRIEF Internal"
df.all[38, 2] = "BRIEF External"
df.all[39, 2] = "BRIEF Total Problems"
df.all[40, 2] = "Social Responsiveness Scale Tot"
df.all[41, 2] = "UPPS-P Negative Urgency"
df.all[42, 2] = "UPPS-P Lack Planning"
df.all[43, 2] = "UPPS-P Positive Urgency"
df.all[44, 2] = "UPPS-P Lack Perserverance"
df.all[45, 2] = "BIS/BAS Reward Responsiveness"
df.all[46, 2] = "BIS/BAS Drive"
df.all[47, 2] = "Parent General Behavior Inventory"
df.all[48, 2] = "Prodromal Psychosis Number"
df.all[49, 2] = "Prodromal Psychosis Severity"

bl <- colorRampPalette(c("navy","royalblue","lightskyblue"))(200)                       
re <- colorRampPalette(c("mistyrose", "red2","darkred"))(200)

plot <- ggplot(df.all[df.all$P_PRS<1,],
               aes(x=reorder(ID, -log10(FDR_PRS) * sign(beta_PRS)), y=-log10(FDR_PRS), fill=beta_PRS)) +
  geom_bar(stat='identity') + coord_flip() + 
  geom_hline(yintercept = -log10(0.05),lty=2) + 
  scale_fill_gradientn(colours=c(bl,"white",re),na.value="grey", limits=c(-1.2,1.2)) +
  theme(legend.position = "right") +
  labs(x="", y="-log(FDR)") +
  ggtitle("MSFc Associations with Mental Health / Cognition")

print(plot)
}
```

## PGS PheWAS - baseline - female #
```{r}
female_phewas <- filter(abcdData_chronotype, sex == "Female")

datalist <- list(female_phewas)
names(datalist) <- c("female_phewas")

# Linear model
for (j in 1:length(datalist)) {
df.linear <- data.frame() 

  for (i in c(345, 346, 349:351, 353, 72, 336, 342)) {
    this_var = datalist[[j]][, i]
    my.model <- lm(as.matrix(this_var) ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                      interview_age + ses_AvgIncomeParentEdu + pds_p_ss_all_category_2 +
                     pc1 + pc2 + pc3 + pc4 +  pc5 + pc6 + pc7 + pc8 + pc9 + pc10 +
                     site_id_l, 
                    data = datalist[[j]], na.action = na.omit)
df.linear <- rbind(df.linear, data.frame(data = names(datalist)[j],
                               ID = colnames(datalist[[j]])[i], 
                               beta_PRS = summary(my.model)$coefficients[2, 1],
                               stderr_PRS = summary(my.model)$coefficients[2, 2],
                               P_PRS = summary(my.model)$coefficients[2, 4]))
  }

# Gamma log-link
df.gamma <- data.frame() 
  for (i in c(10:16, 62:71, 73:81, 314:317, 319, 334:335, 337:338, 2538:2542)) {
    this_var = datalist[[j]][, i]
    my.model <- glm(as.matrix(this_var) ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                      interview_age + ses_AvgIncomeParentEdu + pds_p_ss_all_category_2 +
                      pc1 + pc2 + pc3 + pc4 +  pc5 + pc6 + pc7 + pc8 + pc9 + pc10 + 
                      site_id_l, 
                    data = datalist[[j]], na.action = na.omit, family = Gamma(link = "log"))
df.gamma <- rbind(df.gamma, data.frame(data = names(datalist)[j],
                               ID = colnames(datalist[[j]])[i], 
                               beta_PRS = summary(my.model)$coefficients[2, 1],
                               stderr_PRS = summary(my.model)$coefficients[2, 2],
                               P_PRS = summary(my.model)$coefficients[2, 4]))
  }
  
df.female <- rbind(df.linear, df.gamma)
df.female$FDR_PRS <- p.adjust(df.female$P_PRS,'fdr')
print(df.female)

df.female[1, 2] = "NIH Toolbox Picture Vocabulary"
df.female[2, 2] = "NIH Toolbox Flanker"
df.female[3, 2] = "NIH Toolbox Pattern Comparison"
df.female[4, 2] = "NIH Toolbox Picture Memory"
df.female[5, 2] = "NIH Toolbox Reading"
df.female[6, 2] = "NIH Toolbox Crystalized"
df.female[7, 2] = "CBCL Total Problems"
df.female[8, 2] = "UPPS Sensation Seeking"
df.female[9, 2] = "BIS/BAS Fun Seeking"
df.female[10, 2] = "SDS Maintaining Sleep"
df.female[11, 2] = "SDS Sleep Breathing"
df.female[12, 2] = "SDS Disorders of Arousal"
df.female[13, 2] = "SDS Sleep-Wake Transition Disorders"
df.female[14, 2] = "SDS Excessive Somnolemce"
df.female[15, 2] = "SDS Sleep Hyperhydrosis"
df.female[16, 2] = "SDS Total Sleep Problems"
df.female[17, 2] = "CBCL Anx/Depress"
df.female[18, 2] = "CBCL Withdrawn/Depress"
df.female[19, 2] = "CBCL Somatic Complaints"
df.female[20, 2] = "CBCL Social Problems"
df.female[21, 2] = "CBCL Thought Problems"
df.female[22, 2] = "CBCL Attention Problems"
df.female[23, 2] = "CBCL Rule Breaking"
df.female[24, 2] = "CBCL Aggressive Behavior"
df.female[25, 2] = "CBCL Internalizing"
df.female[26, 2] = "CBCL Externalizing"
df.female[27, 2] = "CBCL Depression"
df.female[28, 2] = "CBCL Anxiety Disorders"
df.female[29, 2] = "CBCL Somatic Problems"
df.female[30, 2] = "CBCL ADHD"
df.female[31, 2] = "CBCL Oppositional Defiant"
df.female[32, 2] = "CBCL Conduct Problems"
df.female[33, 2] = "CBCL Sluggish Cog Tempp"
df.female[34, 2] = "CBCL OCD"
df.female[35, 2] = "CBCL Stress"
df.female[36, 2] = "BREIF Attention"
df.female[37, 2] = "BRIEF Internal"
df.female[38, 2] = "BRIEF External"
df.female[39, 2] = "BRIEF Total Problems"
df.female[40, 2] = "Social Responsiveness Scale Tot"
df.female[41, 2] = "UPPS-P Negative Urgency"
df.female[42, 2] = "UPPS-P Lack Planning"
df.female[43, 2] = "UPPS-P Positive Urgency"
df.female[44, 2] = "UPPS-P Lack Perserverance"
df.female[45, 2] = "BIS/BAS Reward Responsiveness"
df.female[46, 2] = "BIS/BAS Drive"
df.female[47, 2] = "Parent General Behavior Inventory"
df.female[48, 2] = "Prodromal Psychosis Number"
df.female[49, 2] = "Prodromal Psychosis Severity"

plot <- ggplot(df.female[df.female$P_PRS<1,],
               aes(x=reorder(ID, -log10(FDR_PRS) * sign(beta_PRS)), y=-log10(FDR_PRS), fill=beta_PRS)) +
  geom_bar(stat='identity') + coord_flip() + 
  geom_hline(yintercept = -log10(0.05),lty=2) + 
  scale_fill_gradientn(colours=c(bl,"white",re),na.value="grey", limits=c(-1.2,1.2)) +
  theme(legend.position = "right") +
  labs(x="", y="-log(FDR)") +
  ggtitle("Female: Chronotype PGS Associations")

print(plot)
  
}
```

## PRS PheWAS - baseline - male #
```{r}
male_phewas <- filter(abcdData_chronotype, sex == "Male")

datalist <- list(male_phewas)
names(datalist) <- c("male_phewas")

# Linear model
for (j in 1:length(datalist)) {
df.linear <- data.frame() 

  for (i in c(345, 346, 349:351, 353, 72, 336, 342)) {
    this_var = datalist[[j]][, i]
    my.model <- lm(as.matrix(this_var) ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                      interview_age + ses_AvgIncomeParentEdu + pds_p_ss_all_category_2 +
                     pc1 + pc2 + pc3 + pc4 +  pc5 + pc6 + pc7 + pc8 + pc9 + pc10 +
                     site_id_l, 
                    data = datalist[[j]], na.action = na.omit)
df.linear <- rbind(df.linear, data.frame(data = names(datalist)[j],
                               ID = colnames(datalist[[j]])[i], 
                               beta_PRS = summary(my.model)$coefficients[2, 1],
                               stderr_PRS = summary(my.model)$coefficients[2, 2],
                               P_PRS = summary(my.model)$coefficients[2, 4]))
  }

# Gamma log-link
df.gamma <- data.frame() 
  for (i in c(10:16, 62:71, 73:81, 314:317, 319, 334:335, 337:338, 2538:2542)) {
    this_var = datalist[[j]][, i]
    my.model <- glm(as.matrix(this_var) ~ scale(PRS.CHRONOTYPE.JONES.NoMHC, center = TRUE, scale = TRUE) + 
                      interview_age + ses_AvgIncomeParentEdu + pds_p_ss_all_category_2 +
                      pc1 + pc2 + pc3 + pc4 +  pc5 + pc6 + pc7 + pc8 + pc9 + pc10 + 
                      site_id_l, 
                    data = datalist[[j]], na.action = na.omit, family = Gamma(link = "log"))
df.gamma <- rbind(df.gamma, data.frame(data = names(datalist)[j],
                               ID = colnames(datalist[[j]])[i], 
                               beta_PRS = summary(my.model)$coefficients[2, 1],
                               stderr_PRS = summary(my.model)$coefficients[2, 2],
                               P_PRS = summary(my.model)$coefficients[2, 4]))
  }
  
df.male <- rbind(df.linear, df.gamma)
df.male$FDR_PRS <- p.adjust(df.male$P_PRS,'fdr')
print(df.male)

df.male[1, 2] = "NIH Toolbox Picture Vocabulary"
df.male[2, 2] = "NIH Toolbox Flanker"
df.male[3, 2] = "NIH Toolbox Pattern Comparison"
df.male[4, 2] = "NIH Toolbox Picture Memory"
df.male[5, 2] = "NIH Toolbox Reading"
df.male[6, 2] = "NIH Toolbox Crystalized"
df.male[7, 2] = "CBCL Total Problems"
df.male[8, 2] = "UPPS Sensation Seeking"
df.male[9, 2] = "BIS/BAS Fun Seeking"
df.male[10, 2] = "SDS Maintaining Sleep"
df.male[11, 2] = "SDS Sleep Breathing"
df.male[12, 2] = "SDS Disorders of Arousal"
df.male[13, 2] = "SDS Sleep-Wake Transition Disorders"
df.male[14, 2] = "SDS Excessive Somnolemce"
df.male[15, 2] = "SDS Sleep Hyperhydrosis"
df.male[16, 2] = "SDS Total Sleep Problems"
df.male[17, 2] = "CBCL Anx/Depress"
df.male[18, 2] = "CBCL Withdrawn/Depress"
df.male[19, 2] = "CBCL Somatic Complaints"
df.male[20, 2] = "CBCL Social Problems"
df.male[21, 2] = "CBCL Thought Problems"
df.male[22, 2] = "CBCL Attention Problems"
df.male[23, 2] = "CBCL Rule Breaking"
df.male[24, 2] = "CBCL Aggressive Behavior"
df.male[25, 2] = "CBCL Internalizing"
df.male[26, 2] = "CBCL Externalizing"
df.male[27, 2] = "CBCL Depression"
df.male[28, 2] = "CBCL Anxiety Disorders"
df.male[29, 2] = "CBCL Somatic Problems"
df.male[30, 2] = "CBCL ADHD"
df.male[31, 2] = "CBCL Oppositional Defiant"
df.male[32, 2] = "CBCL Conduct Problems"
df.male[33, 2] = "CBCL Sluggish Cog Tempp"
df.male[34, 2] = "CBCL OCD"
df.male[35, 2] = "CBCL Stress"
df.male[36, 2] = "BREIF Attention"
df.male[37, 2] = "BRIEF Internal"
df.male[38, 2] = "BRIEF External"
df.male[39, 2] = "BRIEF Total Problems"
df.male[40, 2] = "Social Responsiveness Scale Tot"
df.male[41, 2] = "UPPS-P Negative Urgency"
df.male[42, 2] = "UPPS-P Lack Planning"
df.male[43, 2] = "UPPS-P Positive Urgency"
df.male[44, 2] = "UPPS-P Lack Perserverance"
df.male[45, 2] = "BIS/BAS Reward Responsiveness"
df.male[46, 2] = "BIS/BAS Drive"
df.male[47, 2] = "Parent General Behavior Inventory"
df.male[48, 2] = "Prodromal Psychosis Number"
df.male[49, 2] = "Prodromal Psychosis Severity"

plot <- ggplot(df.male[df.male$P_PRS<1,],
               aes(x=reorder(ID, -log10(FDR_PRS) * sign(beta_PRS)), y=-log10(FDR_PRS), fill=beta_PRS)) +
  geom_bar(stat='identity') + coord_flip() + 
  geom_hline(yintercept = -log10(0.05),lty=2) + 
  scale_fill_gradientn(colours=c(bl,"white",re),na.value="grey", limits=c(-1.2,1.2)) +
  theme(legend.position = "right") +
  labs(x="", y="-log(FDR)") +
  ggtitle("Male: Chronotype PGS Associations")

print(plot)
  
}
```

## Combine plots #
```{r}
# Merge data frames
df.pgs.all <- rbind(df.female, df.male)

df.pgs.all$neg10_log_fdr <- -log10(df.pgs.all$FDR_PRS)

df.pgs.all$neg10_log_fdr[df.pgs.all$data == "female_phewas"] <-  -df.pgs.all$neg10_log_fdr[df.pgs.all$data == "female_phewas"]

x <- (-log10(0.05))
neg_x <- -x

df.pgs.all %>%
  ggplot() +
  geom_col(aes(x =reorder(ID, abs(neg10_log_fdr)), y =neg10_log_fdr, fill = beta_PRS)) + 
  scale_fill_gradientn(colours=c(bl,"white",re),na.value="grey", limits=c(-1.2,1.2)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.4)) +
  geom_hline(yintercept = 0, lty = 1) +
  geom_hline(yintercept = x, lty = 2) +
  geom_hline(yintercept = neg_x, lty = 2) +
  theme(legend.position = "right") +
  theme(axis.text.x=element_text(hjust=0.95))
```
